### 2021-11-01 text/outline/draft.tex
update abstract
clarify that both Gasp and Xie used thresholding

\begin{abstract}
CRISPR genome engineering and single-cell sequencing have transformed biological discovery. Single-cell CRISPR screens unite these two technologies, linking genetic perturbations in individual cells to changes in gene expression. Single-cell CRISPR screens promise to illuminate the regulatory wiring underlying diseases but present substantial statistical challenges. We demonstrate on real data that a standard method for estimation and inference in single-cell CRISPR screens — “thresholded regression” — exhibits attenuation bias and a bias-variance tradeoff as a function of an intrinsic tuning parameter. We recover these phenomena in precise theoretical terms in an idealized Gaussian setting. Next, we introduce GLM-EIV (“generalized linear model with errors-in-variables”), a new method for single-cell CRISPR screen analysis. GLM-EIV generalizes the classical errors-in-variables model to response distributions and sources of measurement error that are exponential family-distributed, overcoming limitations of thresholded regression. We develop a computational infrastructure to deploy GLM-EIV across hundreds or thousands of processors on clouds (e.g., Microsoft Azure) and high-performance clusters. Leveraging this infrastructure, we apply GLM-EIV to analyze two recent, large-scale, single-cell CRISPR screen datasets, yielding new biological insights.
\end{abstract}

\tableofcontents

\section{Introduction}
CRISPR is a genome engineering tool that has enabled scientists to precisely edit human and nonhuman genomes, opening the door to new medical therapies \cite{Rothgangl2021,Musunuru2021} and transforming basic biology research \cite{Przybyla2021}. Recently, scientists have paired CRISPR genome engineering with single-cell sequencing \cite{Dixit2016,Datlinger2017}. The resulting assays, known as a ``single-cell CRISPR screens,'' link genetic perturbations in individual cells to changes in gene expression, illuminating regulatory networks underlying human diseases and other traits \cite{Morris2021a}.

Despite their promise, single-cell CRISPR screens present substantial statistical challenges. A major difficulty is that CRISPR perturbations are unobservable and assigned stochastically to cells. As a consequence, one cannot know with certainty which cells were perturbed. Instead, one must leverage an indirect, noisy proxy of perturbation presence or absence -- namely, transcribed guide RNA counts -- to ``guess'' which cells were perturbed. Using these imputed perturbation assignments, one can attempt to estimate the effect of the perturbation on gene expression. The standard approach, which we call ``thresholded regression'' or the ``thresholding method,'' is to assign perturbation identities to cells by simply thresholding the guide RNA counts.

@@

Our work makes two main contributions. First, we study the thresholding method from empirical and theoretical perspectives. Notably, we demonstrate on real data that the thresholding method exhibits attenuation bias and a bias-variance tradeoff as a function of the selected threshold, and we recover these phenomena in precise mathematical terms in an idealized Gaussian model. Second, we introduce a new method for estimation and inference in single-cell CRISPR screens that accounts for the measurement error inherent in the experiment. The method, called \textit{GLM-EIV} (generalized linear model with errors-in-variables), implicitly estimates the probability that each cell was perturbed, obviating the need to explicitly impute perturbation assignments via thresholding or another heuristic. Theoretical analyses and simulation studies indicate that GLM-EIV outperforms the thresholding method in large regions of the parameter space.

@@

The genomics literature has produced several applied methods for single-cell CRISPR screen analysis. In a prior work we developed SCEPTRE \cite{Barry2020}, a custom implementation of the conditional randomization test \cite{Candes2018, Liu2021} tailored to single-cell CRISPR screen data. SCEPTRE tests whether a given perturbation is associated with the change in expression of a given gene, adjusting for sources of confounding and ensuring robustness to expression model misspecification. Other applied methods for single-cell CRISPR screen analysis include MIMOSCA \cite{Dixit2016} and scMAGeCK \cite{Yang2019}. These methods, like SCEPTRE, focus on hypothesis testing (rather than estimation), but unlike SCEPTRE, they ignore the countedness of the data and are unable to handle confounders. In this work we tackle a set of analysis challenges that are complimentary to the challenges addressed by SCEPTRE. Most importantly, we seek to \textit{estimate} (with confidence) the effect size of a perturbation on gene expression change, moving beyond hypothesis testing.

@@

The most promising biotechnology for solving this challenge are high MOI single-cell CRISPR screens. High MOI single-cell CRISPR screens combine CRISPR interference (CRISPRi) -- a version of CRISPR that \textit{represses} a targeted region of the genome, leaving the primary DNA sequence intact  -- with single-cell sequencing. The experimental protocol is as follows. First, the scientist develops a library of several hundred to several thousand CRISPRi perturbations, each designed to target a candidate enhancer for repression. The scientist then cultures tens or hundreds of thousands of cells and delivers the CRISPRi perturbations to the cells. The perturbations assort into the cells randomly, with each cell receiving on average 10-40 distinct perturbations. Conversely, a given perturbation enters about 0.1-2\% of cells.

After waiting several days for CRISPRi to take effect, the scientist profiles each cell's transcriptome (i.e., its gene expressions) and the set of perturbations that it received. Finally, the scientist conducts perturbation-to-gene association analyses. Figure \ref{analysis_challenges}a depicts this process schematically, with colored bars (blue, red, and purple) representing distinct perturbations. For a given perturbation (e.g., the perturbation represented in blue), the scientist partitions the cells into two groups: those that received the perturbation (top) and those that did not (bottom). Next, for a given gene, the scientist runs a differential expression analysis across the two groups of cells, producing an estimate for the extent to which the gene's expression has changed in response to the perturbation. If the estimated change in expression is large, the scientist can conclude that the enhancer \textit{targeted} by the perturbation exerts a strong regulatory effect on the gene. This procedure is repeated for a large set of preselected perturbation-gene pairs. % If the differential expression analyses produce $p$-values, the $p$-values are subjected to a multiple hypothesis testing correction procedure.

\subsection{Analysis challenges}
High MOI single-cell CRISPR screens present several statistical challenges. We take a parametric approach in this work, and so we highlight challenges that are especially relevant to modeling the data generating process. Throughout, we consider a single perturbation-gene pair. First, the ``treatment'' variable -- i.e., the presence or absence of a perturbation -- is unobservable. Perturbed cells transcribe molecules called  \textit{guide RNAs} (or \textit{gRNAs}) that serve as indirect signals of perturbation presence. We can leverage the gRNA data to infer whether a given cell has been perturbed. Using these imputed perturbation assignments, we can estimate the effect of the perturbation on gene expression (Figure \ref{analysis_challenges}b). Second, ``technical factors'' -- sources of variation that are experimental rather than biological in origin -- impact both gene expression and gRNA expression (Figure \ref{analysis_challenges}b).  We must account for these technical factors in constructing models for gene and gRNA expression. Third, the gene and gRNA data are sparse, discrete counts. Therefore, classical statistical approaches that assume Gaussianity or homoscedasticity are inapplicable. Finally, and most subtly, sequenced gRNAs sometimes are mapped to unperturbed cells. This phenomenon, which we call the ``background read'' phenomenon, results from errors in the sequencing process \cite{Replogle2020}. The marginal distribution of the gRNA count data is best conceptualized as a mixture model (Figure \ref{analysis_challenges}c; Gaussian distributions used for illustration purposes only). Unperturbed and perturbed cells both exhibit nonzero gRNA count distributions, but the gRNA count distribution for perturbed cells is greater. Figure \ref{analysis_challenges}d shows example data for four (of possibly tens or hundreds of thousands of) cells. The analysis objective is to leverage the observed data -- i.e., the gene expressions and gRNA counts -- to estimate the effect of the (latent) perturbation on gene expression, accounting for the technical factors.

In this work we analyze two large-scale, high MOI, single-cell CRISPRi screen datasets published by Gasperini et al.\ and Xie et al. Gasperini (resp., Xie) targeted approximately 6,000 (resp., 500) candidate enhancers in a population of approximately 200,000 (resp., 100,000) cells. Gasperini additionally designed 381 positive control, gene-targeting perturbations and 50 non-targeting, negative control perturbations to assess method sensitivity and specificity.

\section{Thresholding method}

The method that Gasperini et al.\ used in their original analysis is the so-called ``thresholding method,'' or ``thresholded regression.'' In this section we study the thresholding method from empirical and theoretical perspectives, uncovering several basic limitations of the approach. Let $n \in \N$ be the number of cells assayed in the experiment. Consider a single perturbation and a single gene. For cell $i \in \{1, \dots, n\}$, let $m_i \in \N$ be the number of gene transcripts sequenced; let $g_i \in \N$ be the number of gRNA transcripts sequenced; let $l^m_i \in \N$ be the number of gene transcripts sequenced across \textit{all} genes (the library size); and finally, let $z_i \in \R^{d-1}$ be the cell-specific technical factors (e.g., sequencing batch, percent mitochondrial reads, etc.) The letters ``m,'' ``g'', and ``l'' stand for ``mRNA,'' ``gRNA,'' and ``library,'' respectively. The thresholding method is defined as follows:
\begin{itemize}
\item[1.] For a given threshold $c \in \N$, let the imputed perturbation assignment $\hat{p}_i \in \{0, 1\}$ be $$\begin{cases} \hat{p}_i = 0 \textrm{ if } g_i \geq c, \\ \hat{p}_i = 1 \textrm{ if } g_i < c. \end{cases}$$
\item[2.] Assume that $m_i$ is related to $\hat{p}_i, l_i,$ and $z_i$ through the following GLM:
\begin{equation*}
\begin{cases}
m_i | (\hat{p}_i, z_i, l_i) \sim \textrm{NB}_\theta(\mu_i), \\
\log(\mu_i) = \beta^m_0 + \beta^m_1 \hat{p}_i + \gamma^T_m z_i + \log\left( l_i^m\right),
\end{cases}
\end{equation*}
where (i) $NB_\theta(\mu_i)$ is a negative binomial distribution with mean $\mu_i$ and known size parameter $\theta$; (ii) $\beta^m_0 \in \R, \beta^m_1 \in \R,$ and $\gamma \in  \R^{d-1}$ are unknown parameters; and (iii) $\log(l_i^m)$ is an offset term. Fit a GLM to obtain estimates of the parameters.

@@

The biological interpretation for the target of inference $\beta^m_1$ as follows:  $\beta^m_1$ is the log-transformed fold change in gene expression in response to the perturbation, removing the effect of the technical factors. Fold change (obtained by exponentiating $\beta^m_1$) is the ratio of the mean gene expression in cells that were perturbed to the mean gene expression in cells that were not perturbed. $\exp(\beta^m_1) = 1$ indicates no change in mean expression, while $\exp(\beta^m_1) > 1$ and $\exp(\beta^m_1) < 1$ indicate an increase and decrease in expression in response to the perturbation, respectively (accounting for technical factors).

@@

We hypothesized, however, that the reduction in bias conferred by selecting a large threshold comes with the cost of increasing the variance of the estimator. To investigate, we compared $p$-values and confidence intervals produced by threshold = 5 and threshold = 20 for the target of inference $\beta^m_1$. We found that threshold = 5 yielded smaller (i.e., more significant) $p$-values and narrower confidence intervals than did threshold = 20 (Figure \ref{thresholding_empirical}c-d). We concluded that the threshold controls a bias-variance tradeoff: as the threshold increases, bias decreases and variance increases.

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=1\linewidth]{../../figures/thresholding_empirical/plot}
	\caption{\textbf{Empirical challenges of thresholded regression.} \textbf{a-b,} Fold change estimates produced by threshold = 1 versus threshold = 5 (a) and threshold = 20 versus threshold = 5 (b). The selected threshold substantially impacts the results. \textbf{c-d,} $p$-values (c) and CI widths (d) produced by threshold = 20 versus threshold = 5. The latter threshold yields more confident estimates. \textbf{e-f}, Empirical distribution of randomly-selected gRNA from Gasperini (e) and Xie (f) data. The gRNA data do not appear to imply an obvious threshold selection strategy.}
	\label{thresholding_empirical}
\end{figure}
\newpage

@@

Next, we study the thresholding method from a theoretical perspective, recovering the attenuation bias and bias-variance tradeoff effects established in the previous section, as well as several other interesting phenomena.  We work in an idealized Gaussian setting. Suppose that we observe gRNA and gene expression data $\{(g_1, m_1), \dots, (g_n, m_n)\}$ on $n \in \N$ cells from the following model:

@@

Empirical and theoretical analyses reveal that the thresholding method poses several challenges: the threshold is a tuning parameter that substantially impacts the results; strict attenuation bias obtains uniformly over the parameter space and for all choices of the threshold; and there does not exist and obvious threshold selection strategy due to (i) the unimodality of the empirical gRNA count distributions and (ii) the existence of a bias-variance tradeoff mediated by the threshold. These difficulties motivate our core research question: \textit{Does modeling the gRNA count distribution directly, thereby circumventing the need to threshold altogether, lead to simpler, more accurate estimation and inference in single-cell CRISPR screen analysis?} To answer this question, we introduce a generalization of the classical errors-in-variables model to response distributions and sources of measurement error that are exponential family-distributed.

>>>>

\begin{abstract}
CRISPR genome engineering and single-cell sequencing have transformed biological discovery. Single-cell CRISPR screens unite these two technologies, linking genetic perturbations in individual cells to changes in gene expression. Single-cell CRISPR screens promise to shed light on the genetic architecture underlying diseases but present substantial statistical challenges. %  In this work we study single-cell CRISPR screens from a statistical perspective.
We demonstrate on real data that a standard method for estimation and inference in single-cell CRISPR screens — “thresholded regression” — exhibits attenuation bias and a bias-variance tradeoff as a function of an intrinsic tuning parameter. We recover these phenomena in precise theoretical terms in an idealized Gaussian setting. Next, we introduce GLM-EIV (“generalized linear model with errors-in-variables”), a new method for single-cell CRISPR screen analysis. GLM-EIV generalizes the classical errors-in-variables model to response distributions and sources of measurement error that are exponential family-distributed, overcoming limitations of thresholded regression. We develop a computational infrastructure to deploy GLM-EIV across hundreds or thousands of processors on clouds (e.g., Microsoft Azure) and high-performance clusters. Leveraging this infrastructure, we apply GLM-EIV to analyze two recent, large-scale, single-cell CRISPR screen datasets, yielding new biological insights.
\end{abstract}
\tableofcontents
\noindent \blue{All edits in blue.}
\section{Introduction}
CRISPR is a genome engineering tool that has enabled scientists to precisely edit human and nonhuman genomes, opening the door to new medical therapies \cite{Rothgangl2021,Musunuru2021} and transforming basic biology research \cite{Przybyla2021}. Recently, scientists have paired CRISPR genome engineering with single-cell sequencing \cite{Dixit2016,Datlinger2017}. The resulting assays, known as a ``single-cell CRISPR screens,'' link genetic perturbations in individual cells to changes in gene expression, illuminating regulatory networks underlying human diseases and other traits \cite{Morris2021a}.

Despite their promise, single-cell CRISPR screens present substantial statistical challenges. A major difficulty is that CRISPR perturbations \blue{are assigned stochastically to cells and cannot be observed directly}. As a consequence, one cannot know with certainty which cells were perturbed. Instead, one must leverage an indirect, noisy proxy of perturbation presence or absence -- namely, transcribed guide RNA counts -- to ``guess'' which cells were perturbed. Using these imputed perturbation assignments, one can attempt to estimate the effect of the perturbation on gene expression. The standard approach, which we call ``thresholded regression'' or the ``thresholding method,'' is to assign perturbation identities to cells by simply thresholding the guide RNA counts.

@@

Our work makes two main contributions. First, we study the thresholding method from empirical and theoretical perspectives. Notably, we demonstrate on real data that the thresholding method exhibits attenuation bias and a bias-variance tradeoff as a function of the selected threshold, and we recover these phenomena in precise mathematical terms in an idealized Gaussian model. \blue{Second, we introduce a new method, GLM-EIV (generalized linear model with errors-in-variables), for single-cell CRISPR screen analysis. GLM-EIV generalizes the classical errors-in-variables model to response distributions and sources of measurement error that are exponential family-distributed. GLM-EIV implicitly estimates the probability that each cell was perturbed, obviating the need to explicitly impute perturbation assignments via thresholding or another heuristic.} Theoretical analyses and simulation studies indicate that GLM-EIV outperforms the thresholding method in large regions of the parameter space.

@@

The genomics literature has produced several applied methods for single-cell CRISPR screen analysis. In a prior work we developed SCEPTRE \cite{Barry2020}, a custom implementation of the conditional randomization test \cite{Candes2018, Liu2021} tailored to single-cell CRISPR screen data. SCEPTRE tests whether a given perturbation is associated with the change in expression of a given gene, adjusting for sources of confounding and ensuring robustness to expression model misspecification. Other applied methods for single-cell CRISPR screen analysis include MIMOSCA \cite{Dixit2016} and scMAGeCK \cite{Yang2019}. These methods, like SCEPTRE, focus on hypothesis testing (rather than estimation), but unlike SCEPTRE, they ignore the countedness of the data and are unable to handle confounders. In this work we tackle a set of analysis challenges that are complimentary to the challenges addressed by SCEPTRE. Most importantly, we seek to \textit{estimate} (with confidence) the effect size of a perturbation on gene expression change, \blue{a statistical objective that is unattainable within the nonparameteric hypothesis testing framework of SCEPTRE.}

@@

The most promising biotechnology for solving this challenge are high MOI single-cell CRISPR screens. High MOI single-cell CRISPR screens combine CRISPR interference (CRISPRi) -- a version of CRISPR that represses a targeted region of the genome -- with single-cell sequencing. The experimental protocol is as follows. First, the scientist develops a library of several hundred to several thousand CRISPRi perturbations, each designed to target a candidate enhancer for repression. The scientist then cultures tens or hundreds of thousands of cells and delivers the CRISPRi perturbations to these cells. The perturbations assort into the cells randomly, with each cell receiving on average 10-40 distinct perturbations. Conversely, a given perturbation enters about 0.1-2\% of cells.

After waiting several days for CRISPRi to take effect, the scientist profiles each cell's transcriptome (i.e., its gene expressions) and the set of perturbations that it received. Finally, the scientist conducts perturbation-to-gene association analyses. Figure \ref{analysis_challenges}a depicts this process schematically, with colored bars (blue, red, and purple) representing distinct perturbations. For a given perturbation (e.g., the perturbation represented in blue), the scientist partitions the cells into two groups: those that received the perturbation (top) and those that did not (bottom). Next, for a given gene, the scientist runs a differential expression analysis across the two groups of cells, producing an estimate for the magnitude of the gene expression change in response to the perturbation. If the estimated change in expression is large, the scientist can conclude that the enhancer \textit{targeted} by the perturbation exerts a strong regulatory effect on the gene. This procedure is repeated for a large set of preselected perturbation-gene pairs. \blue{The enhancer-by-enhancer approach is valid because the perturbations assort into cells approximately independently of one another.}

\subsection{Analysis challenges}
\blue{High MOI single-cell CRISPR screens present several statistical challenges, four of which we highlight here.} Throughout, we consider a single perturbation-gene pair. First, the ``treatment'' variable -- i.e., the presence or absence of a perturbation -- cannot be directly observed. Instead, perturbed cells transcribe molecules called  \textit{guide RNAs} (or \textit{gRNAs}) that serve as indirect proxies of perturbation presence. We must leverage these gRNAs to impute (explicitly or implicitly) perturbation assignments onto the cells (Figure \ref{analysis_challenges}b). \blue{Second, ``technical factors'' -- sources of variation that are experimental rather than biological in origin -- impact both gene expression and gRNA expression (Figure \ref{analysis_challenges}b). Technical factors act as confounders in the measurement process and therefore must be accounted for in differential expression models.} Third, the gene and gRNA data are sparse, discrete counts. Therefore, classical statistical approaches that assume Gaussianity or homoscedasticity are inapplicable. Finally, and most subtly, sequenced gRNAs sometimes are mapped to cells that have not received a perturbation. This phenomenon, which we call the ``background read'' phenomenon, results from errors in the sequencing and alignment processes \cite{Replogle2020}. The marginal distribution of the gRNA counts is best conceptualized as a mixture model (Figure \ref{analysis_challenges}c; Gaussian distributions used for illustration purposes only). \blue{Unperturbed and perturbed cells both exhibit nonzero gRNA count distributions, but this distribution overall is greater for perturbed cells.} Figure \ref{analysis_challenges}d shows example data on four (of possibly tens or hundreds of thousands of) cells. \blue{The analysis objective is to leverage the gene expressions and gRNA counts to estimate the effect of the (latent) perturbation on gene expression, accounting for the technical factors.}

In this work we analyze two large-scale, high MOI, single-cell CRISPR screen datasets published by Gasperini et al.\ and Xie et al. Gasperini (resp., Xie) targeted approximately 6,000 (resp., 500) candidate enhancers in a population of approximately 200,000 (resp., 100,000) cells. Gasperini additionally designed 381 positive control, gene-targeting perturbations and 50 non-targeting, negative control perturbations to assess method sensitivity and specificity.

\section{Thresholding method}
\blue{In their original analysis of the data, Gasperini and Xie imputed perturbation identities onto the cells by thresholding the gRNA counts. In this section we study the thresholding method from empirical and theoretical perspectives, uncovering several limitations of the approach. We investigate the specific version of the thresholding method used by Gasperini, which involved a thresholding step and a follow-up negative binomial regression step. (Xie implemented a slightly different and, in our view, less principled version of thresholding, which we do not review here.)}

Let $n \in \N$ be the number of cells assayed in the experiment. Consider a single perturbation and a single gene. For cell $i \in \{1, \dots, n\}$, let $m_i \in \N$ be the number of gene transcripts sequenced; let $g_i \in \N$ be the number of gRNA transcripts sequenced; let $l^m_i \in \N$ be the number of gene transcripts sequenced across \textit{all} genes (the library size); and finally, let $z_i \in \R^{d-1}$ be the cell-specific technical factors (e.g., sequencing batch, percent mitochondrial reads, etc.) The letters ``m,'' ``g'', and ``l'' stand for ``mRNA,'' ``gRNA,'' and ``library,'' respectively. The thresholding method is defined as follows:
\begin{itemize}
\item[1.] For a given threshold $c \in \N$, let the imputed perturbation assignment $\hat{p}_i \in \{0, 1\}$ be $$\begin{cases} \hat{p}_i = 0 \textrm{ if } g_i < c, \\ \hat{p}_i = 1 \textrm{ if } g_i \geq c. \end{cases}$$
\item[2.] Assume that $m_i$ is related to $\hat{p}_i, l_i,$ and $z_i$ through the following GLM:
$$m_i | (\hat{p}_i, z_i, l_i) \sim \textrm{NB}_\theta(\mu_i),$$
\begin{equation}\label{thresh_glm}
\log(\mu_i) = \beta^m_0 + \beta^m_1 \hat{p}_i + \gamma^T_m z_i + \log\left( l_i^m\right),
\end{equation}
where (i) $NB_\theta(\mu_i)$ is a negative binomial distribution with mean $\mu_i$ and known size parameter $\theta$; (ii) $\beta^m_0 \in \R, \beta^m_1 \in \R,$ and $\gamma \in  \R^{d-1}$ are unknown parameters; and (iii) $\log(l_i^m)$ is an offset term. Fit a GLM to obtain estimates of the parameters.

@@

% Setting $\gamma_m^T$ to zero and exponentiating both sides of (\ref{thresh_glm}), we obtain
% $$\mu_i = \left(e^{\beta^m_0 + \beta^m_1\hat{p}_i}\right) l^m_i.$$ The mean expression of unperturbed cells $\mu_i(0)$ is
% $$\mu_i(0) = e^{\beta^m_0} l_i^m,$$ and the mean expression in perturbed cells is
% $$ \mu_i(1) = e^{\beta^m_1} e^{\beta^m_0} l^m_i.$$ Therefore,
% $$ \beta^m_1 = \log\left(\frac{\mu_i(1)}{\mu_i(0)}\right).$$
% We use library size $l_i^m$

The biological interpretation for the target of inference $\beta^m_1$ as follows:  $\beta^m_1$ is the log-transformed fold change in gene expression in response to the perturbation, \blue{controlling for} the technical factors. Fold change (obtained by exponentiating $\beta^m_1$) is the ratio of the mean gene expression in cells that were perturbed to the mean gene expression in cells that were not perturbed. $\exp(\beta^m_1) = 1$ indicates no change in mean expression, while $\exp(\beta^m_1) > 1$ and $\exp(\beta^m_1) < 1$ indicate an increase and decrease in expression in response to the perturbation, respectively (accounting for technical factors).

@@

We hypothesized, however, that the reduction in bias conferred by selecting a large threshold comes at the cost of increasing the variance of the estimator. To investigate, we compared $p$-values and confidence intervals produced by threshold = 5 and threshold = 20 for the target of inference $\beta^m_1$. We found that threshold = 5 yielded smaller (i.e., more significant) $p$-values and narrower confidence intervals than did threshold = 20 (Figure \ref{thresholding_empirical}c-d). We concluded that the threshold controls a bias-variance tradeoff: as the threshold increases, bias of the estimator decreases and variance increases.

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=1\linewidth]{../../figures/thresholding_empirical/plot}
	\caption{\textbf{Empirical challenges of thresholded regression.} \textbf{a-b,} Fold change estimates produced by threshold = 1 versus threshold = 5 (a) and threshold = 20 versus threshold = 5 (b). The selected threshold substantially impacts the results. \textbf{c-d,} $p$-values (c) and CI widths (d) produced by threshold = 20 versus threshold = 5. The latter threshold yields more confident estimates. \textbf{e-f}, Empirical distribution of randomly-selected gRNA from Gasperini (e) and Xie (f) data (0 counts not shown). The gRNA data do not appear to imply an obvious threshold selection strategy.}
	\label{thresholding_empirical}
\end{figure}
\newpage

@@

Next, we study the thresholding method from a theoretical perspective, recovering in precise mathematical terms the attenuation bias and bias-variance tradeoff effects uncovered on real data, as well as several other interesting phenomena.  We work in an idealized Gaussian setting. Suppose that we observe gRNA and gene expression data $\{(g_1, m_1), \dots, (g_n, m_n)\}$ on $n \in \N$ cells from the following model:

@@

Empirical and theoretical analyses reveal that the thresholding method poses several challenges: the threshold is a tuning parameter that substantially impacts the results; strict attenuation bias obtains uniformly over the parameter space and for all choices of the threshold; and there does not exist and obvious threshold selection strategy due to (i) the unimodality of the empirical gRNA count distributions and (ii) the existence of a bias-variance tradeoff mediated by the threshold. These difficulties motivate our core research question: \textit{Does modeling the gRNA count distribution directly, thereby circumventing the need to threshold altogether, lead to simpler, more accurate estimation and inference in single-cell CRISPR screen analysis?} To answer this question, we generalize the classical errors-in-variables model to response distributions and sources of measurement error that are exponential family-distributed.
